The impact of cosmic variance on simulating weak lensing surveys

### 2014-10-07 Writeup/paper_v3.tex
restructure and reorder abstract

One method that is commonly used to test for the presence of
systematic errors in the shear estimation process is image simulation,
where we can cleanly test whether our methods of shear estimation
recover the ground truth. This is a valuable test, considering the
numerous sources of additive and multiplicative bias such as a
mismatch between galaxy model assumptions and actual galaxy light
profiles \citep[e.g.,][]{2010MNRAS.404..458V, 2010A&A...510A..75M}, biases due to the effects of pixel noise on the
shear estimates
\citep{2012MNRAS.427.2711K,2012MNRAS.424.2757M,2012MNRAS.425.1951R},
and ellipticity gradients \citep{2010MNRAS.406.2793B}.  These biases
often differ for galaxies with different morphologies (e.g., disks
vs. ellipticals), sizes, $S/N$, and shape.  
 Realistic simulations use samples based on images from the Hubble
 Space Telescope ({\em HST}) such as COSMOS \citep[used by the GREAT3 challenge,][]{great3} or the
 Ultra Deep Field \citep[UDF, used by][]{2013ApJ...765...74J}.  These
 two examples serve as the extremes in the {\em HST} samples used as
 the basis for image simulation, with the former being shallower but
 the current widest contiguous area surveyed by the {\em HST}, and the
 latter being extremely deep but narrow.
The goal of this work is to understand the systematic uncertainties
that might arise due to the finite size of these areas, resulting in
cosmic variance that could result in selection of atypical galaxy
samples, particularly when dividing the sample as a function of
redshift. Upcoming surveys will study lensing as a function of
redshift and therefore need to simulate galaxy samples at different
redshifts in order to assess the shear calibration at each redshift.

%In almost all branches of observational cosmology, weak lensing introduces systematic error, in addition to itself being a probe of the dark matter and eventually the large scale structure of the universe.

A general requirement for simulations used to test shear recovery is
that they should be as realistic as possible.  
With this in mind, the GREAT3 challenge~\citep{great3} was held with the aim to test methods for measuring weak gravitational lensing by analyzing astronomical images.
Software packages like
GalSim\footnote{\url{https://github.com/GalSim-developers/GalSim}}
\citep{2014arXiv1407.7676R} can generate images of galaxies from the
{\em HST} as they would appear with an additional lensing shear and
viewed by some lower resolution telescope.
The training dataset used in the challenge comes from the COSMOS
survey, which we will describe in \S\ref{S:data}.
%While providing a good quality (high $S/N$) images to be trained over, there could be some bias introduced due to the survey being a narrow one.
For a variety of physical reasons, some of which are still not fully
understood, the shape and morphology of
galaxies depends on their local environment
\citep[e.g.,][]{2014arXiv1402.1172C,2014MNRAS.444.2200D}. 
Hence, local overdensities or underdensities
observed in the COSMOS field may (given the small size of the field)
cause the properties of the galaxy population in redshift slices to be
atypical depending on the environment in that slice.  This would have
the undesired consequence of including variation in galaxy properties
due to the COSMOS galaxy sample in the simulated galaxies, rather than
only including ensemble effects such as true redshift evolution of
galaxy properties.  Our goal is to quantify the degree to which the
morphology-density correlations in COSMOS cause noticeable changes in
the galaxy populations in narrow redshift slices at a level that could
result in difficulty using the sample to derive redshift-dependent
shear calibrations.

>>>>

One method that is commonly used to test for the presence of
systematic errors in the shear estimation process is image simulation,
where we can cleanly test whether our methods of shear estimation
recover the ground truth. This is a valuable test, considering the
numerous sources of additive and multiplicative bias such as a
mismatch between galaxy model assumptions and actual galaxy light
profiles \citep[e.g.,][]{2010MNRAS.404..458V, 2010A&A...510A..75M}, biases due to the effects of pixel noise on the
shear estimates
\citep{2012MNRAS.427.2711K,2012MNRAS.424.2757M,2012MNRAS.425.1951R},
and ellipticity gradients \citep{2010MNRAS.406.2793B}.  These biases
often differ for galaxies with different morphologies (e.g., disks
vs. ellipticals), sizes, $S/N$, and shape \citep{2010MNRAS.405.2044B,2012MNRAS.423.3163K}. 
A general requirement for simulations used to test shear recovery is
that they should be as realistic as possible.

 Realistic simulations may use samples based on images from the Hubble
 Space Telescope ({\em HST}).
Software packages like
GalSim\footnote{\url{https://github.com/GalSim-developers/GalSim}}
\citep{2014arXiv1407.7676R} can generate images of galaxies from the
{\em HST} as they would appear with an additional lensing shear and
viewed by some lower resolution telescope.  Examples of training samples from the {\em HST} include 
the COSMOS survey \citep[used by the GREAT3 challenge,][]{great3} or the
 Ultra Deep Field \citep[UDF, used by][]{2013ApJ...765...74J}.  These
 two examples serve as the extremes in the {\em HST} samples used as
 the basis for image simulation, with COSMOS being shallower but
representing the current widest contiguous area surveyed by the {\em HST}, and the
 latter being extremely deep but narrow.

For a variety of physical reasons, some of which are still not fully
understood, the shape and morphology of
galaxies depends on their local environment
\citep[e.g.,][]{2014arXiv1402.1172C,2014MNRAS.444.2200D}. 
Hence, local overdensities or underdensities
observed in these {\em HST} fields may (given the small size of the field)
cause the properties of the galaxy population in redshift slices to be
atypical depending on the environment in that slice.  This would have
the undesired consequence of including variation in galaxy properties
due to the COSMOS (or other) survey cosmic variance in the simulated galaxy sample in that redshift slice, rather than
only including ensemble effects that would appear in a large cosmological volume, such as true redshift evolution of
galaxy properties.  Our goal is to quantify the degree to which the
morphology-density correlations in COSMOS cause noticeable changes in
the galaxy populations in narrow redshift slices at a level that could
result in difficulty using the sample to derive redshift-dependent
shear calibrations.   Upcoming surveys will study lensing as a function of
redshift and therefore need to simulate galaxy samples at different
redshifts in order to assess the shear calibration at each redshift.
